\documentclass{article}
\usepackage{amsmath}
\usepackage[margin=1.5cm]{geometry}
\usepackage{graphicx}
\begin{document}

\begin{center}
	{\huge \textbf{Part 4: Statistics}}

	Calculators are allowed in this section
\end{center}

\bigskip

\section{Probability}

\bigskip

\textbf{Give the probability of the event occurring}

\bigskip

	\indent A bag contains six red marbles and five blue marbles. Another bag contains seven green marbles and three yellow marbles. You randomly pick one marble from each bag. One marble is blue and one marble is yellow.

	\bigskip

	\bigskip

	\indent There are eight nickels and five dimes in your pocket. You randomly pick a coin out of your pocket and place it on the counter. Then you pick another. Both coins are nickels.

	\bigskip

	\bigskip

	\indent A  bag contains six red marbles and eight blue marbles. Another bag contains five green marbles and eight yellow marbles. You randomly pick one marble from each bag. One marble is blue and one marble is yellow.

	\bigskip

	\bigskip

	\indent You flip a coin twice. Both flips land tails-up.

	\bigskip

	\bigskip

	\indent A box of chocolates contains four milk hocolates, five dark chocolates, and three white chocolates. You randomly select a chocolate. It is a milk chocolate or a dark chocolate.

	\bigskip

	\bigskip

	\indent A basket contains five apples, four peaches, and five pears. You randomly select a piece of fruit. It is an apple or a peach.

	\bigskip

	\bigskip

	\indent A litter of kittens consists of one gray female, two gray males, three black females, and one black male.
You randomly pick one kitten. The kitten is black or female.
	
	\bigskip

	\bigskip

	\indent There are ten shirts in your closet. Six are blue and four are green. Three of the blue shirts and four of the green shirts fit well. The others are too big. You randomly select a shirt to wear. It is green and fits well.

	\bigskip

	\bigskip



\bigskip
\section{Mean, Median, and Mean Absolute Deviation}
\bigskip
\textbf{Find the mean and median, and Mean absolute deviation of the following sets}
\bigskip

\(\begin{aligned}
	&{47,27,9,15,15,19} \qquad\qquad {9,23,38,45,14} \qquad\qquad {37,27,1,31,27,40,47} \qquad\qquad {9,43,50,0}\\\\
	&{40,8,31,3} \qquad\qquad {19,10,0,1,35} \qquad\qquad {5,42,24,43,34} \qquad\qquad {5,42,24,43,34} \qquad\qquad {38,16,37,29}\\\\
	&{16,27,24,13,15,2,50} \qquad\qquad {48,26,6,26} \qquad\qquad {29,39,28,50} \qquad\qquad {27,46,11,12,35}\\\\
\end{aligned}\)

\end{document}
